---
title: "Grossepieper_Nils_Assignment02"
author: "Nils Grossepieper"
date: "2023-03-27"
output: html_document
---

```{r setup, include = F}
options(scipen=999)
```

***

### Happiness data

<div class="alert alert-block alert-info">
(@) Begin by importing the data on happiness. Sadly, the data is only available in several separate data sets. Import the data on the years 2015 through 2019 and merge them. You are going to need to identify and fix problems with incompatible column names.
</div>

```{r libraries, warning=FALSE}

# Load all important libraries:
library(tidyverse)
library(ggplot2)

```
```{r}

# Set up the working directory:
path <- "C:/Users/gross/OneDrive/Dokumente/Grossepieper_Nils_Assignment02"
setwd(path)

# Import the data sets for the years 2015 up to 2019:
happy_2015 <- read_csv("happy_data/happy_2015.csv", na = c("", "NA"))
happy_2016 <- read_csv("happy_data/happy_2016.csv", na = c("", "NA"))
happy_2017 <- read_csv("happy_data/happy_2017.csv", na = c("", "NA"))
happy_2018 <- read_csv("happy_data/happy_2018.csv", na = c("", "NA"))
happy_2019 <- read_csv("happy_data/happy_2019.csv", na = c("", "NA"))

# Add a year variable to every data set:
happy_2015$Year <- replicate(nrow(happy_2015), 2015)
happy_2016$Year <- replicate(nrow(happy_2016), 2016)
happy_2017$Year <- replicate(nrow(happy_2017), 2017)
happy_2018$Year <- replicate(nrow(happy_2018), 2018)
happy_2019$Year <- replicate(nrow(happy_2019), 2019)

# Set up subsets of the data frames containing only the important variables:
happy_2015 <- happy_2015 %>% select(Year, Country, `Happiness Rank`, `Happiness Score`)
happy_2016 <- happy_2016 %>% select(Year, Country, `Happiness Rank`, `Happiness Score`)
happy_2017 <- happy_2017 %>% select(Year, Country, Happiness.Rank, Happiness.Score)
happy_2018 <- happy_2018 %>% select(Year, `Country or region`, `Overall rank`, Score)
happy_2019 <- happy_2019 %>% select(Year, `Country or region`, `Overall rank`, Score)

# Rename all columns based on the happy_2017 variable names:
happy_2015 <- happy_2015 %>% rename("Happiness.Rank" = 3,
                                    "Happiness.Score" = 4)
happy_2016 <- happy_2016 %>% rename("Happiness.Rank" = 3,
                                    "Happiness.Score" = 4)
happy_2018 <- happy_2018 %>% rename("Country" = 2,
                                    "Happiness.Rank" = 3,
                                    "Happiness.Score" = 4)
happy_2019 <- happy_2019 %>% rename("Country" = 2,
                                    "Happiness.Rank" = 3,
                                    "Happiness.Score" = 4)

# Now we can merge the five data sets to one data set happy:
happy <- rbind(happy_2015, happy_2016, happy_2017, happy_2018, happy_2019)
head(happy, 10)

```

***

### Inflation data

Data on inflation is available from World Bank. 

<div class="alert alert-block alert-info">
(@) Download the most current version of the data from [their web site](https://data.worldbank.org/indicator/FP.CPI.TOTL.ZG). 
(@) Import the data calling it `inflation` and turn it into a tidy format.
</div>

```{r}

# Import the inflation data:
inflation <- read_csv("API_FP.CPI.TOTL.ZG_DS2_en_csv_v2_4701153/API_FP.CPI.TOTL.ZG_DS2_en_csv_v2_4701153.csv",
                      skip = 4)

# Set up subsets of the data frames containing only the important variables:
inflation <- inflation %>% select(`Country Name`, `2015`, `2016`, `2017`, `2018`,
                                  `2019`)

# Transform the data set inflation into a tidy data format:
inflation_tidy <- inflation %>% pivot_longer(cols = c(`2015`, `2016`, `2017`,
                                                      `2018`, `2019`), names_to = "Year",
                                             values_to = "Inflation")

# Rename the column Country Name:
inflation_tidy <- inflation_tidy %>% rename("Country" = 1)
head(inflation_tidy, 10)

```

***

### Merging the data

<div class="alert alert-block alert-info">
(@) Merge the two data frames `happy` and `inflation` by country and year. Only keep those observations for which the key exists in both tables.
</div>

```{r}

# Check if the variables have the same type and if not correct this:
glimpse(happy)
glimpse(inflation_tidy)
inflation_tidy$Year <- as.numeric(inflation_tidy$Year)

# Now I perform a inner join on the data sets happy and inflation_tidy:
happy_inflation <- inner_join(happy, inflation_tidy, by = c("Country", "Year"))

# Remove all NA values:
happy_inflation <- na.omit(happy_inflation)
head(happy_inflation, 10)

```

<div class="alert alert-block alert-info">
(@) Which table contains data about more countries than the other? How many of these countries are dropped by merging the data sets?
</div>

```{r}

# Count the unique countries in happy, inflation_tidy and the combination of both
# data sets:
length(unique(happy$Country))
length(unique(inflation_tidy$Country))
sum(unique(happy$Country) %in% unique(inflation_tidy$Country))

# The data set tidy_inflation has more countries. 
# When you merge the two data sets (without dropping the NA-values) you end up 
# with 142 this means that you loose 124 countries from inflation_tidy have been
# dropped

```





